Algebraic Surfaces

### 2016-05-24 project/article_revised_revised/main.tex
final

\input{code_style.tex}
\usepackage[]{microtype} 
\usepackage[]{enumitem} 
\usepackage[]{graphicx} 
\usepackage[]{subcaption} 
\usepackage[]{makeidx} 
\makeindex

@@

\usepackage[noabbrev]{cleveref}

@@

\DeclareMathOperator{\Col}{Col}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\Bl}{Bl}
\title{Algebraic Surfaces}
\author{Ivar Haugal{\o}kken Stangeby}
\date{\textsc{MAT2000 - Spring 2016}}

@@

    \begin{abstract}
       In this paper we give a brief overview of some of the things one can be
       looking at when studying algebraic surfaces.  Mainly, we look at
       singularities of such surfaces, how they arise and how one can remove
       them. The images are rendered using the software \textsc{Surfer} unless
       specified otherwise. Some auxilliary \textsc{Python} and
       \textsc{Mathematica} scripts has also been used.
    \end{abstract}

@@

    \begin{corollary}
        Given a set of polynomials $f_1, f_2, \ldots, f_m$, the variety
        $\V(f_1, f_2, \ldots, f_m)$ can be expressed as a finite intersection
        \begin{equation}
            \notag
            \V(f_1, \ldots, f_m) = \bigcap_{i=1}^{m} \V(f_i).
        \end{equation}
    \end{corollary}
    It is also of interest, especially when aiming to create and visualize
    interesting figures, that the following hold:
    \begin{corollary}
    Given a set of polynomials $f_1, f_2, \ldots, f_m$, the affine algebraic
    variety of the product $f_1f_2\cdots f_m$ is expressible as a finite union
    \begin{equation}
        \notag
        \V(f_1f_2\cdots f_m) = \bigcup_{i=1}^{m} \V(f_i).
    \end{equation}
    \end{corollary}

    With this in mind, we define one of the more interesting properties of
    varieties, and one we will work with a lot in this article, namely the
    singularities. A singular point is, strictly speaking, a point on a surface
    where the behaviour of the surface is slightly strange, say the surface
    passes through itself. We are often interested in studying how a surface
    behaves locally near a point $p$. We do this using \emph{linearization at
    $p$}. This is what we do when we consider the tangent line to a point for a
    single-valued function, the tangent plane for a general surface in three
    dimensions, and a tangent \emph{hyperplane} for general surfaces in $n$
    dimensions. The linearization of a surface given by $f(x_1, x_2, \ldots,
    x_n) = 0$, at a point $p_0$ is given by
    \begin{equation}
        \notag
        L(p) = f(p_0) + J_f(p_0)(p - p_0)
    \end{equation}
    where $J_f(p_0)$ is the Jacobian matrix of $f(x_1, x_2, \ldots, x_n)$
    evaluated at $p_0$. If this matrix has rank less than $n$, then the
    linearization is not uniquely defined, i.e., there are more than one plane
    satisfying the linearization. This leads us to the next defintion.

@@

    \begin{definition}[The projective space of dimension $n$\index{projective space}]
        Let $\sim$ be a relation on $\C^{n+1}\setminus \left\{ 0 \right\}$ defined by
        \begin{equation}
            \notag
            (x_1, \ldots, x_{n+1}) \sim (x_1', \ldots, x_{n+1}')
            \iff (x_1', \ldots, x_{n+1}') = \lambda(x_1, \ldots,
            x_{n+1})
        \end{equation}
        where $\lambda$ is some non-zero complex number. This relation can be
        shown to be an equivalence relation. We denote the
        equivalence class of $(x_1, x_2, \ldots, x_{n+1})$ by $\left[ x_1 : x_2
        : \ldots : x_{n+1} \right]$ and define the \emph{projective space of
        dimension $n$} as the set 
        \begin{equation}
            \notag
            \P^{n}_\C = \left\{ \left[ x_1 : \ldots : x_{n+1} \right] \mid (x_1,
            \ldots , x_{n+1}) \in \C^{n+1}\setminus \left\{ 0 \right\}\right\}.
        \end{equation}
        The elements in $\P_\C^n$ are called \emph{homogeneous
        coordinates}\index{homogeneous coordinates}. 
    \end{definition}

@@

    \section{Introduction to algebraic surfaces}

    In this section, we try to motivate the definition of an algebraic surface,
    and look at some examples. Hopefully, this will give an intuitive
    introduction to exactly what we will we looking at. We start by looking at
    real algebraic curves, as a good entry point.

@@

    From elementary mathematics, one learns about real valued functions,
    $f(x)$, and how to plot these functions by setting $y = f(x)$ and plotting
    points in the $(x, y)$-plane. This yields a curve of sorts, moving around
    in the plane.

    The graph of a function is something of a peculiarity, because it comes
    with some restrictions. Not all curves in the $(x, y)$-plane correspond to
    a specific function. One of the standard methods for verifying whether a
    certain curve corresponds to a function or not, typically taught in school,
    is the \emph{vertical line test}. If you can draw a vertical line through
    two or more points on the curve, then the curve is not associated to any
    specific function. There is however a limit to how interesting
    function-curves can be.

    Having an equation $y = f(x)$ we can form the \emph{equation of a curve at
    zero}. We define a new function in two variables and equate it to zero.
    This yields
    \begin{equation}
        \notag
        g(x, y) = y - f(x) = 0.
    \end{equation}
    The set of points along the curve described by this equation is precisely
    the points that lie in the variety $\V(g(x, y))$.
    If the function $f(x)$ is a polynomial in the variable $x$, i.e., $f(x) \in
    K[x]$, we call the function curve of $f(x)$ algebraic. If the distinction
    between a function graph and curves is to be justified, then there has to
    be curves that are not function graphs.

@@

    \begin{figure}[h!]
        \centering
        \begin{subfigure}[t]{0.45\textwidth}
            \includegraphics[width=\textwidth]{pictures/node.pdf}
            \caption{The curve given by the equation $y^2 - x^3 - x^2 = 0$.
            This curve is commonly known as ``The Node''. It exhibits a
            singularity at the point $p = (0, 0)$ which can be verified by
            computing the partial derivatives at $p$.}
            \label{fig:node}
        \end{subfigure}        
        ~
        \begin{subfigure}[t]{0.45\textwidth}
            \includegraphics[width=\textwidth]{pictures/cusp.pdf}
            \caption{The curve given by $y^2 - x^3 = 0$, or more commonly known
            as ``The Cusp'', is also an example of a singular algebraic curve.
        The point $p = (0, 0)$ is a point of singularity.}
        \end{subfigure}
        \caption{Curves not corresponding to functions.}
        \label{fig:algebraic_curves}
    \end{figure}

    \subsection{Moving up a dimension}
    \label{sub:moving_up_a_dimension}
    With what we have done so far, we can generalize and move up a dimension.
    Instead of considering curves in the $(x, y)$-plane, we can instead look at
    surfaces in the three-dimensional $(x, y, z)$-space. Following the same
    procedure as above, we define a new function $h(x, y, z) = z - g(x, y)$.
    Completely analogous to the definition made for curves, we call the surface
    \emph{algebraic} if $h(x, y, z)$ is a polynomial in the variables $x, y,
    z$. That is $h(x, y, z) \in K[x, y, z]$. We give some examples.

    \begin{example}
        \label{exmpl:algebraic_surfaces}
        The following equations satisfy the definition of an algebraic surface:
        \begin{enumerate}
            \item $h_1(x, y, z) = z - xy = 0$,
            \item $h_2(x, y, z) = z - x^2 - y^2 = 0$,
            \item $h_3(x, y, z) = z - (2x^2 - y)(y - x^2) = 0$,
        \end{enumerate}
        while the following does \emph{not}:
        \begin{enumerate}[resume]
            \item $g_1(x, y, z) = \sin(xy) + z = 0$, and
            \item $g_2(x, y, z) = \tan^2(x) + \pi z y$.
        \end{enumerate}
        These examples are visualized in \cref{fig:algebraic_surfaces}.
        \begin{figure}[]
            \centering
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[height=0.6\textwidth]{pictures/example_one.png} 
                \caption{$h_1(x, y, z) = 0$}
            \end{subfigure}
            ~
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[height=0.6\textwidth]{pictures/example_two.png} 
                \caption{$h_2(x, y, z) = 0$}
            \end{subfigure}
            ~
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[height=0.6\textwidth]{pictures/example_three.png} 
                \caption{$h_3(x, y, z) = 0$}
            \end{subfigure}
            \caption{The surfaces that fell under the category algebraic from
            \cref{exmpl:algebraic_surfaces}.}
            \label{fig:algebraic_surfaces}
        \end{figure}
    \end{example}
    We can of course continue this process of introducing new functions and
    move up a dimension for each time indefinitely.  We complete this section
    by formally defining an algebraic surface. In any case, the surfaces are
    defined by the variety of some defining polynomial.
    \begin{definition}[Algebraic surface\index{algebraic surface}]
        An \emph{algebraic surface} is an algebraic variety of
        dimension\footnote{The dimension of an algebraic variety is a bit hard
            to grasp, and since we only will be dealing with algebraic surfaces
            in this text we do not spend any time classifying the different
        kinds of varieties. The explanation of this is therefore omitted. See
    \cite{Dim16}.}
            two.  We define the \emph{degree}\index{algebraic surface!degree}
            of an algebraic surface to be the highest combined power in the
            defining polynomials terms, and denote this $d$.  That is,
        \begin{equation}
            \notag
            d = \max_{\alpha}\left\{ \alpha_1 + \alpha_2 + \ldots \alpha_n \right\}
        \end{equation}
        where $\alpha_i$ is the power of $x_i$ in a term.
    \end{definition}
    We now take a brief look at some of the properties of these surfaces in
    relation to the singularities.

    \section{Singularities}
    \label{sec:singularities}
    The singularities of an algebraic surface is one of the important elements
    that give the surface its visual appeal. I am sure most people agree with
    me, when I say that a surface full of singularities is more intriguing than
    one with no singularities at all. The singularities, along with their
    configurations, symmetries and their number are things that one can explore
    further.

@@

    From \cite[p.~7]{Lab14} we have the following upper bound on the maximum
    number of singularities for $d \geq 3$,
    \begin{equation}
        \notag
        \mu(d) \leq \frac{1}{2}(d(d-1)^2 - 3)
    \end{equation}
    and plugging in $d = 3$ yields $\mu(d) \leq 4$. However, we have already
    presented a surface of degree $d = 3$ with four singularities, so in fact
    $\mu(d) = 4$. The theory of finding the maximal number of singularities of
    a given degree and constructing surfaces that exhibit all is a very
    technical matter. The thesis \cite{oliver2005hypersurfaces} gives a good
    introduction to the construction and visualization of such surfaces.

@@

    We now turn to the resolution of singularities. There are, as mentioned
    above, many ways of doing this, but we will here apply a geometrical
    approach.

    \subsection{Blowing up a point singularity.}
    \label{sub:blowing_up_a_point_singularity_}

@@

        Setting $v = 1$ yields the equation $x = yu$ which makes us able to
        eliminate $x$ from \cref{eq:node}. This gives us a new equation
        \begin{equation}
            \notag
            h(y, u) = y^2(1 - u^2 - u^2y) = 0.
        \end{equation}
        This is the other affine chart of $\Bl_p\A^2$ and can be visualized in
        the $(y, u)$-plane.  The affine charts given by $h(x, y, z)$ and $g(x,
        y, z)$ are shown visualized in \cref{fig:affine_charts_node}.

        The relation between these two charts, and the original surface can be
        visualized as in \cref{fig:blow_up_intuitive}.

        \begin{figure}[h!]
            \centering
            \begin{subfigure}[t]{0.4\textwidth}
                \includegraphics[width=1.1\textwidth]{pictures/node_affine_1.pdf}
                \caption{The affine chart resulting from setting $u = 1$ and
                eliminating $y$ in the original equation.}
            \end{subfigure}
            ~
            \begin{subfigure}[t]{0.4\textwidth}
                \includegraphics[width=1.1\textwidth]{pictures/node_affine_2.pdf}
                \caption{The affine chart resulting from setting $v = 1$ and
                eliminating $x$ in the original equation}
            \end{subfigure}
            \begin{subfigure}[t]{0.4\textwidth}
                \includegraphics[width=1.0\textwidth]{pictures/blow_up_intuitive.pdf}
                \caption{A visual depiction of the blowup surface that clearly
                    shows how the singularity in the origin is resolved. Taken
                from \cite{Whi08}.}
                \label{fig:blow_up_intuitive}
            \end{subfigure}
            \caption{The affine charts of the Node. The vertical line plotted
                in slightly brighter blue, is called the \emph{exceptional
                divisor}\index{exceptional divisor} and corresponds to the
                blowup of what was the origin in the original surface. Note
                that the curves intersect the exceptional divisor in exactly
                two points. This corresponds to the double point singularity in the
                original surface.}
            \label{fig:affine_charts_node}
        \end{figure}
    \end{example}

@@

    \subsection{Blowing up a line of singularity.}
    \label{sub:blowing_up_a_line_of_singularity_}

@@

    \begin{figure}[h]
        \centering
        \includegraphics[width=0.4\linewidth]{pictures/blowup_line.png}
        \caption{The ``Whitney Umbrella'' defined by the polynomial $f(x, y, z)
        = x^2y - z^2$. This surface exhibits singularities all along the
        $y$-axis, as can easily be pointed out on the picture.}
        \label{fig:whitney_umbrella}
    \end{figure}

@@

        \begin{figure}[h]
            \centering
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[width=\textwidth]{pictures/blowup_line_2.png}
                \caption{$g(x, s, t) = 0$}
            \end{subfigure}      
            ~
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[width=\textwidth]{pictures/blowup_line_1.png}
                \caption{$h(y, r, t) = 0$}
                \label{fig:whitney_blowup}
            \end{subfigure}      
            ~
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[width=\textwidth]{pictures/blowup_line_3.png}
                \caption{$i(z, r, s) = 0$}
            \end{subfigure}      
        \caption{The affine charts of the blowup of the Whitney Umbrella at the
            point $p = (0, 0, 0)$. As we can see the singularities improve in
            two of the charts, and in one case the chart is smooth. However,
            the middle surface is identical to our original surface, so the
            blow-up has failed to resolve all the singularities of the original
            variety.}
            \label{fig:blowup_line_charts}
        \end{figure}      
    \end{example}

@@

    \begin{definition}[Blowup of $\A^3$ at a line $L$]
        We let the blowup-surface $\Bl_L\A^3$ be the set of all pairs
        consisting of a point $p \in \A^3$ and a point $\lambda \in \P^1$ such
        that $H_\lambda$ is the unique plane that contains both the point $p$
        and the line $L$ and write
        \begin{equation}
            \notag
            \Bl_L\A^3 = \left\{ (p, \lambda) \in \A^3 \times \P^1 \mid p, L \in H_\lambda \right\}.
        \end{equation}
    \end{definition}

    We illustrate this definition by again considering the Whitney Umbrella.
    \begin{example}[Blowup of the Whitney Umbrella at a line $L$]
        The singularities of the Whitney Umbrella lie along the $y$-axis, which
        is described by the equations $x = z = 0$. Now, our goal here is to
        find under what circumstances a plane $H_\lambda$ contains the point $p
        = (x, y, z)$ as well as the line $L$. In other words, we want both the
        $y$-axis and $p$ to lie in the plane $rx + sy + tz = 0$.

        We first find the planes that contain the $y$-axis. Choosing the vector
        $(0, 1, 0)$ as a representative for the $y$-axis we compute the cross
        product between this vector and the point $(r, s, t)$ as a
        representative for the plane.  This yields a normal vector given by
        \begin{equation}
            \notag
            \vec{n} = \begin{vmatrix}
                \hat{i} & \hat{j} & \hat{k} \\
                r & s & t \\
                0 & 1 & 0
            \end{vmatrix} = (-t, 0, r).
        \end{equation}
        A plane containing the $y$-axis is then given by the equation $-tx + 0y
        + rz = rz - tx = 0$. We can now consider the two affine charts by
        setting $t = 1$ and $s = 1$ respectively.

>>>>

\usepackage[]{fontspec} 
\input{code_style.tex}
\usepackage[]{microtype} 
\usepackage[]{enumitem} 
\usepackage[]{graphicx} 
\usepackage[]{subcaption} 
\usepackage[]{makeidx}

\usepackage[margin=2cm]{geometry} 
\makeindex

@@

@@

\DeclareMathOperator{\Col}{Col}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\Bl}{Bl}

\usepackage[]{hyperref} 
\usepackage[noabbrev]{cleveref} 
\title{Algebraic Surfaces}
\author{Ivar Haugal{\o}kken Stangeby}
\date{\textsc{MAT2000 - Spring 2016}}

@@

    \begin{abstract}
       In this paper we give a brief overview of some of the things one can be
       looking at when studying algebraic surfaces. We use a fairly naive
       approach without the very abstract mathematical machinery that this
       field has to offer. In \cref{sec:preliminaries} we give some of the
       needed mathematical preliminaries. \Cref{sec:intro} motivates the
       definition of an algebraic surface, the main object of study. In
       \cref{sec:singularities} we look at one of the main properties of
       algebraic surfaces, namely the singular points, and proceed by showing
       how to resolve them i \cref{sec:resolving_singularities}. Finally, we
       give an example of how to construct a singular surface with a certain
       symmetry in \cref{sec:constructing_singular_surfaces}.  The images are
       rendered using the software \textsc{Surfer} unless specified otherwise.
       Some auxiliary \textsc{Python} and \textsc{Mathematica} scripts has also
       been used. 
    \end{abstract}

@@

    \begin{corollary}
        \label{crl:set_one}
        Given a set of polynomials $f_1, f_2, \ldots, f_m$, the variety
        $\V(f_1, f_2, \ldots, f_m)$ can be expressed as a finite intersection
        \begin{equation}
            \notag
            \V(f_1, \ldots, f_m) = \bigcap_{i=1}^{m} \V(f_i).
        \end{equation}
    \end{corollary}
    It is also of interest, especially when aiming to create and visualize
    interesting figures, that the following hold:
    \begin{corollary}
        \label{crl:set_two}
        Given a set of polynomials $f_1, f_2, \ldots, f_m$, the affine algebraic
        variety of the product $f_1f_2\cdots f_m$ is expressible as a finite union
        \begin{equation}
            \notag
            \V(f_1f_2\cdots f_m) = \bigcup_{i=1}^{m} \V(f_i).
        \end{equation}
    \end{corollary}

    With this in mind, we define one of the more interesting properties of
    varieties, and one we will work with a lot in this article, namely the
    singularities. A singular point is, strictly speaking, a point on a surface
    where the behaviour of the surface is slightly strange, say the surface
    passes through itself. We are often interested in studying how a surface
    behaves locally near a point $p$. We do this using \emph{linearization at
    $p$}. This is what we do when we consider the tangent line to a point for a
    single-valued function, the tangent plane for a general surface in three
    dimensions, and a tangent \emph{hyperplane} for general surfaces in $n$
    dimensions. The linearization of a surface given by $f(x_1, x_2, \ldots,
    x_n) = 0$, at a point $p_0$ is given by
    \begin{equation}
        \notag
        L(p) = f(p_0) + J_f(p_0)(p - p_0)
    \end{equation}
    where $J_f(p_0)$ is the Jacobian matrix of $f(x_1, x_2, \ldots, x_n)$
    evaluated at $p_0$. If this matrix has rank less than $n$, then the
    linearization is not uniquely defined, i.e., there are more than one plane
    satisfying the linearization. This leads us to the next definition.

@@

    \begin{definition}[The projective space of dimension $n$\index{projective space}]
        Let $\sim$ be a relation on $\C^{n+1} \setminus \left\{ 0 \right\}$ defined by
        \begin{equation}
            \notag
            (x_1, \ldots, x_{n+1}) \sim (x_1', \ldots, x_{n+1}')
            \iff (x_1', \ldots, x_{n+1}') = \lambda(x_1, \ldots,
            x_{n+1})
        \end{equation}
        where $\lambda$ is some non-zero complex number. This relation can be
        shown to be an equivalence relation. We denote the
        equivalence class of $(x_1, x_2, \ldots, x_{n+1})$ by $\left[ x_1 : x_2
        : \ldots : x_{n+1} \right]$ and define the \emph{projective space of
        dimension $n$} as the set 
        \begin{equation}
            \notag
            \P^{n}_\C = \left\{ \left[ x_1 : \ldots : x_{n+1} \right] \mid (x_1,
            \ldots , x_{n+1}) \in \C^{n+1}\setminus \left\{ 0 \right\}\right\}.
        \end{equation}
        The elements in $\P_\C^n$ are called \emph{homogeneous
        coordinates}\index{homogeneous coordinates}. 
    \end{definition}

@@

    \section{Introduction to algebraic surfaces}
    \label{sec:intro}    
    In this section, we try to motivate the definition of an algebraic surface,
    and look at some examples. Hopefully, this will give an intuitive
    introduction to exactly what we will we looking at. We start by looking at
    real algebraic curves, as a good entry point.

@@

    From elementary mathematics, one learns about real valued functions,
    $f(x)$, and how to plot these functions by setting $y = f(x)$ and plotting
    points in the $(x, y)$-plane. This yields a curve of sorts, moving around
    in the plane.  Having an equation $y = f(x)$ we can form the \emph{equation
    of a curve at zero}. We define a new function in two variables and equate
    it to zero. This yields
    \begin{equation}
        \notag
        g(x, y) = y - f(x) = 0.
    \end{equation}
    The set of points along the curve described by this equation is precisely
    the points that lie in the variety $\V(g(x, y))$.  If the function $f(x)$
    is a polynomial in the variable $x$, i.e., $f(x) \in K[x]$, we call the
    function curve of $f(x)$ algebraic. Now, all the curves formed in this way
    corresponds to some function. There are however curves that does not
    correspond to functions and these are given as general polynomials.

@@

    \begin{figure}[h!]
        \centering
        \begin{subfigure}[t]{0.45\textwidth}
            \includegraphics[width=\textwidth]{pictures/node.pdf}
            \caption{The curve given by the equation $y^2 - x^3 - x^2 = 0$.
            This curve is commonly known as ``The Node''. It exhibits a
            singularity at the point $p = (0, 0)$ which can be verified by
            computing the partial derivatives at $p$.}
            \label{fig:node}
        \end{subfigure}        
        ~
        \begin{subfigure}[t]{0.45\textwidth}
            \includegraphics[width=\textwidth]{pictures/cusp.pdf}
            \caption{The curve given by $y^2 - x^3 = 0$, or more commonly known
            as ``The Cusp'', is also an example of a singular algebraic curve.
        The point $p = (0, 0)$ is a point of singularity.}
        \end{subfigure}
        \caption{First examples of algebraic curves. Visualized using
        \textsc{Python}. Code can be seen in \cref{sec:code_snippets}.}
        \label{fig:algebraic_curves}
    \end{figure}

    \subsection{Moving up a dimension}
    \label{sub:moving_up_a_dimension}
    With what we have done so far, we can generalize and move up a dimension.
    Instead of considering curves in the $(x, y)$-plane, we can instead look at
    surfaces in the three-dimensional $(x, y, z)$-space. Following the same
    procedure as above, we define a new function $h(x, y, z) = z - g(x, y)$.
    This function equated at zero yields a surface.  Completely analogous to
    the definition made for curves, we call the surface \emph{algebraic} if
    $h(x, y, z)$ is a polynomial in the variables $x, y, z$. That is $h(x, y,
    z) \in K[x, y, z]$. We give some examples.

    \begin{example}
        \label{exmpl:algebraic_surfaces}
        The following equations satisfy the definition of an algebraic surface:
        \begin{enumerate}
            \item $h_1(x, y, z) = z - xy = 0$,
            \item $h_2(x, y, z) = z - x^2 - y^2 = 0$,
            \item $h_3(x, y, z) = z - (2x^2 - y)(y - x^2) = 0$,
        \end{enumerate}
        while the following does \emph{not}:
        \begin{enumerate}[resume]
            \item $g_1(x, y, z) = \sin(xy) + z = 0$, and
            \item $g_2(x, y, z) = \tan^2(x) + \pi z y$.
        \end{enumerate}
        These examples are visualized in \cref{fig:algebraic_surfaces}.
        \begin{figure}[]
            \centering
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[height=0.6\textwidth]{pictures/example_one.png} 
                \caption{$h_1(x, y, z) = 0$}
            \end{subfigure}
            ~
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[height=0.6\textwidth]{pictures/example_two.png} 
                \caption{$h_2(x, y, z) = 0$}
            \end{subfigure}
            ~
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[height=0.6\textwidth]{pictures/example_three.png} 
                \caption{$h_3(x, y, z) = 0$}
            \end{subfigure}
            \caption{The surfaces that fell under the category algebraic from
            \cref{exmpl:algebraic_surfaces}.}
            \label{fig:algebraic_surfaces}
        \end{figure}
    \end{example}
    We can of course continue this process of introducing new functions and
    move up a dimension for each time indefinitely.  We complete this section
    by formally defining an algebraic surface. In any case, the surfaces are
    defined by the variety of some defining polynomial.
    \begin{definition}[Algebraic surface\index{algebraic surface}]
        Given a set of polynomials $f_1, f_2, \ldots, f_m$ in three variables,
        i.e., in $K[x, y, z]$ we define an \emph{algebraic surface} to be the
        algebraic variety $\V(f_1, f_2, \ldots, f_m)$.  We define the
        \emph{degree}\index{algebraic surface!degree} of an algebraic surface
        to be the highest combined power in the defining polynomials terms, and
        denote this $d$. That is,
        \begin{equation}
            \notag
            d = \max_{\alpha}\left\{ \alpha_1 + \alpha_2 + \ldots \alpha_n \right\}
        \end{equation}
        where $\alpha_i$ is the power of $x_i$ in a term.
    \end{definition}
    We now take a brief look at some of the properties of these surfaces in
    relation to the singularities.

    \section{Singularities}
    \label{sec:singularities}

    When studying algebraic surfaces we typically differentiate between two
    types of surfaces, namely the singular ones and the non-singular ones.
    These two categories are each interesting in their own respect. A lot of
    mathematical theory concerning surfaces only hold for surfaces that exhibit
    no singularities at all. It is therefore often of interest to somehow
    remove these singularities from the surface in order to apply known
    results.

    The theory of singular surfaces on the other hand contains a lot of
    interesting theory that deals with: the classification of singularities,
    what different types of singularities are there; resolution of
    singularities; and deformation theory.

@@

    From \cite[p.~7]{Lab14} we have the following upper bound on the maximum
    number of singularities for $d \geq 3$,
    \begin{equation}
        \notag
        \mu(d) \leq \frac{1}{2}(d(d-1)^2 - 3)
    \end{equation}
    and plugging in $d = 3$ yields $\mu(3) \leq 4$. However, we have already
    presented a surface of degree $d = 3$ with four singularities, so in fact
    $\mu(d) = 4$. The theory of finding the maximal number of singularities of
    a given degree and constructing surfaces that exhibit all is a very
    technical matter. The thesis \cite{oliver2005hypersurfaces} gives a good
    introduction to the construction and visualization of such surfaces.

@@

    When we blow up, we blow up along a \emph{center} of the variety. By making
    sure that this center is contained in the singular locus of the variety, we
    do not change the surrounding surface. We first consider the case where the
    blowup center is a single point, and then an example where the center is a
    line.

    \subsection{Blowing up a point}
    \label{sub:blowing_up_a_point_singularity_}

@@

        Setting $v = 1$ yields the equation $x = yu$ which makes us able to
        eliminate $x$ from \cref{eq:node}. This gives us a new equation
        \begin{equation}
            \notag
            h(y, u) = y^2(1 - u^2 - u^2y) = 0.
        \end{equation}
        This is the other affine chart of $\Bl_p\A^2$ and can be visualized in
        the $(y, u)$-plane.  The affine charts given by $h(x, y, z)$ and $g(x,
        y, z)$ are shown visualized in \cref{fig:affine_charts_node}.  The
        relation between these two charts, and the original surface can be
        visualized as in \cref{fig:blow_up_intuitive}.

        \begin{figure}[h!]
            \centering
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[width=1.1\textwidth]{pictures/node_affine_1.pdf}
                \caption{The affine chart resulting from setting $u = 1$ and
                eliminating $y$ in the original equation.}
            \end{subfigure}
            ~
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[width=1.1\textwidth]{pictures/node_affine_2.pdf}
                \caption{The affine chart resulting from setting $v = 1$ and
                eliminating $x$ in the original equation}
            \end{subfigure} \\
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[width=1.0\textwidth]{pictures/blow_up_intuitive.pdf}
                \caption{A visual depiction of the blowup surface that clearly
                    shows how the singularity in the origin is resolved. Taken
                from \cite{Whi08}.}
                \label{fig:blow_up_intuitive}
            \end{subfigure}
            \caption{The affine charts of the Node. The vertical line plotted
                in slightly brighter blue, is called the \emph{exceptional
                divisor}\index{exceptional divisor} and corresponds to the
                blowup of what was the origin in the original surface. Note
                that the curves intersect the exceptional divisor in exactly two
                points. This corresponds to the double point singularity in the
                original surface. The top row is visualized in \textsc{Python}.}
            \label{fig:affine_charts_node}
        \end{figure}
    \end{example}

@@

    \subsection{Blowing up a line}
    \label{sub:blowing_up_a_line_of_singularity_}

@@

    \begin{figure}[h]
        \centering
        \includegraphics[width=0.3\linewidth]{pictures/blowup_line.png}
        \caption{The ``Whitney Umbrella'' defined by the polynomial $f(x, y, z)
        = x^2y - z^2$. This surface exhibits singularities all along the
        $y$-axis, as can easily be pointed out on the picture.}
        \label{fig:whitney_umbrella}
    \end{figure}

@@

        \begin{figure}[h]
            \centering
            \begin{subfigure}[t]{0.2\textwidth}
                \includegraphics[width=\textwidth]{pictures/blowup_line_2.png}
                \caption{$g(x, s, t) = 0$}
            \end{subfigure}      
            ~
            \begin{subfigure}[t]{0.2\textwidth}
                \includegraphics[width=\textwidth]{pictures/blowup_line_1.png}
                \caption{$h(y, r, t) = 0$}
                \label{fig:whitney_blowup}
            \end{subfigure}      
            ~
            \begin{subfigure}[t]{0.2\textwidth}
                \includegraphics[width=\textwidth]{pictures/blowup_line_3.png}
                \caption{$i(z, r, s) = 0$}
            \end{subfigure}      
        \caption{The affine charts of the blowup of the Whitney Umbrella at the
            point $p = (0, 0, 0)$. As we can see the singularities improve in
            two of the charts, and in one case the chart is smooth. However,
            the middle surface is identical to our original surface, so the
            blow-up has failed to resolve all the singularities of the original
            variety.}
            \label{fig:blowup_line_charts}
        \end{figure}      
    \end{example}

@@

    \begin{definition}[Blowup of $\A^3$ at a line $L$] 
        Given a point $p$ and a line $L$, we are interested in the unique plane
        $H$ that contains both $p$ and $L$.  Since it contains $L$, the plane
        can be parametrized by a point $\lambda \in \P^1$.

        We let the blowup-surface $\Bl_L\A^3$ be the set of all pairs
        consisting of a point $p \in \A^3$ and a point $\lambda \in \P^1$ that
        parametrizes the unique plane $H$ that contains both $p$ and $L$.  We
        write
        \begin{equation}
            \notag
            \Bl_L\A^3 = \left\{ (p, \lambda) \in \A^3 \times \P^1 \mid p, L \in H \right\}.
        \end{equation}
    \end{definition}

    Again, this is a fairly technical definition, so where do we go from here?
    Assuming $L$ is parametrized by $x = at, y = bt, z = ct$, with $a, b$ and
    $c$ non-zero, we can form the symmetric equations 
    \begin{equation}
        \notag
        t = \frac{x}{a} = \frac{y}{b} = \frac{z}{c}.
    \end{equation}
    The line $L$ is then completely determined by two of the three possible
    relations $xb - ya = 0$, $xc - za = 0$ and $yc - zb = 0$.  In order to find
    the respective plane, we are interested in knowing when the plane $rx + sy
    + tz = 0$ can be expressed in terms of the line given by $xb - ya = 0$ and
    $yc - zb = 0$. This can be solved as a matrix problem. Given the matrix
    \begin{equation}
        \notag
        A = \begin{bmatrix}
            b & 0 & r \\
            -a & c & s \\
            0 & -b & t
        \end{bmatrix}
    \end{equation}
    one of the columns of $A$ is linearly dependent of the two others if
    $\det(A) = 0$. Solving $\det(A) = 0$ gives the equation
    \begin{equation}
        \notag
        \det(A) = b(ct + sb) + rab = 0.
    \end{equation}
    We can now solve for either one of $r, s$ and $t$. Solving for $r$ we get
    $r = -( ct + sb ) / a$ which we can substitute back into the equation for
    our plane.  The plane is then determined by 
    \begin{equation}
        \notag
        \frac{-ct - sb}{a}x + sy + tz = 0.
    \end{equation}
    Multiplying through by $a$ we achieve
    \begin{align*}
                                t(za - cx) + s(ya - bx) = 0,
    \end{align*}
    which gives us the restrictions that we need. The $\lambda$ that
    parametrizes the plane $H$ is then the point $[t:s] \in \P^1$.

    We illustrate this definition by again considering the Whitney Umbrella.
    \begin{example}[Blowup of the Whitney Umbrella at a line $L$]
        The singularities of the Whitney Umbrella lie along the $y$-axis, which
        is described by the equations $x = z = 0$. Now, our goal here is to
        find under what circumstances a plane $H$ contains the point $p = (x,
        y, z)$ as well as the line $L$. In other words, we want both the
        $y$-axis and $p$ to lie in the plane $rx + sy + tz = 0$.

        We first find the planes that contain the $y$-axis. Choosing the vector
        $(0, 1, 0)$ as a representative for the $y$-axis we compute the cross
        product between this vector and the point $(r, s, t)$ as a
        representative for the plane. This yields a normal vector given by
        \begin{equation}
            \notag
            \vec{n} = \begin{vmatrix}
                \hat{i} & \hat{j} & \hat{k} \\
                r & s & t \\
                0 & 1 & 0
            \end{vmatrix} = (-t, 0, r).
        \end{equation}
        A plane containing the $y$-axis is then given by the equation $-tx + 0y
        + rz = rz - tx = 0$. The point $\lambda$ is in this case the point $[r:
        t] \in \P^1$. We can now consider the two affine charts by setting $r =
        1$ and $t = 1$ respectively.
